Collocated Snapshot Backup with VM-Centric Deduplication in the Cloud

### 2013-09-24 snapshot_system/fast14_architecture.tex
revisions

To read a non-PDS chunk using its reference with container ID and local chunk ID,  the snapshot store client first loads the
corresponding VM's container index file specified by the container ID, then searches the chunk
groups  using their  chunk ID coverage.
After that, it reads the identified chunk group from DFS, decompresses it, and seeks to the exact chunk data 
specified by the chunk ID. 
Finally, the client updates its internal chunk data cache with the newly loaded content to 
anticipate future sequential reads.
\item {\em Delete()}.
Chunk deletion occurs when a snapshot expires or gets deleted explicitly by a user
and we will discuss the snapshot deletion in detail in the following subsection.
When deletion requests are issued for a specific container,
those requests are simply recorded into the  container's deletion log initially and thus  a lazy
deletion strategy is exercised.
Once local chunk IDs appear in
the deletion log, they will not be referenced by any future snapshot and can be safely deleted when needed. 
Periodically, the snapshot  store picks those containers with an excessive
number of deletion requests to  compact and  reclaim the corresponding disk space. 
During compaction, the snapshot store creates a new container (with the same container ID) to replace the 
existing one. This is done by sequentially scanning the old container, copying all the chunks that are not 
found in the deletion log to the new container, and creating new chunk groups and indices. 
Every local chunk ID however is directly copied rather than re-generated. This
process leaves holes in the CID values, but preserves the order and IDs of chunks.
As a result, all data references stored 
in recipes are permanent and stable, and the data reading process
is as efficient as before. Maintaining the stability of chunk IDs also ensures that recipes do not
depend directly on physical storage locations, which simplifies data migration.
\end{itemize}

@@

\item {\bf Periodic repair of leakage}.
%[exlpain why second Bloom filter, why scan append store]
Leakage repair is conducted periodically to fix the above approximation error.
This procedure compares the live chunks for each VM with what are truly used in the VM snapshot recipes.
That requires a scanning of all chunks in a VM. Since it is a VM-specific procedure 
the space and time cost is relatively small compared to the traditional mark-sweep procedure~\cite{Guo2011} which scans snapshot 
chunks from all VMs.
For example,
consider each reference consumes 8 bytes plus  1 mark bit, a VM that has 40GB backup data with about
10 million chunks will need less than 90MB of memory to complete a VM-specific mark-sweep process.
\end{itemize}

>>>>

To read a non-PDS chunk using its reference with container ID and local chunk ID,  the snapshot store client first loads the
corresponding VM's container index file specified by the container ID, then searches the chunk
groups  using their  chunk ID coverage.
After that, it reads the identified chunk group from DFS, decompresses it, and seeks to the exact chunk data 
specified by the chunk ID. 
Finally, the client updates its internal chunk data cache with the newly loaded content to 
anticipate future sequential reads.
\item {\em Delete()}.
Chunk deletion occurs when a snapshot expires or gets deleted explicitly by a user
and we will discuss the snapshot deletion in detail in the following subsection.
When deletion requests are issued for a specific container,
those requests are simply recorded into the  container's deletion log initially and thus  a lazy
deletion strategy is exercised.
Once local chunk IDs appear in
the deletion log, they will not be referenced by any future snapshot and can be safely deleted when needed. 
This is ensured because we only dedup against the direct parent of a snapshot, so the deleted snapshot's blocks
will only be used if they also exist in other snapshots.
Periodically, the snapshot  store picks those containers with an excessive
number of deletion requests to  compact and  reclaim the corresponding disk space. 
During compaction, the snapshot store creates a new container (with the same container ID) to replace the 
existing one. This is done by sequentially scanning the old container, copying all the chunks that are not 
found in the deletion log to the new container, and creating new chunk groups and indices. 
Every local chunk ID however is directly copied rather than re-generated. This
process leaves holes in the CID values, but preserves the order and IDs of chunks.
As a result, all data references stored 
in recipes are permanent and stable, and the data reading process
is as efficient as before. Maintaining the stability of chunk IDs also ensures that recipes do not
depend directly on physical storage locations, which simplifies data migration.
\end{itemize}

@@

\item {\bf Periodic repair of leakage}.
%[exlpain why second Bloom filter, why scan append store]
Leakage repair is conducted periodically to fix the above approximation error.
This procedure compares the live chunks for each VM with what are truly used in the VM snapshot recipes.
That requires a scanning of all chunks in a VM. Since it is a VM-specific procedure 
the space and time cost is relatively small compared to the traditional mark-sweep procedure~\cite{Guo2011} which scans snapshot 
chunks from all VMs.
For example,
consider each reference consumes 8 bytes plus  1 mark bit, a VM that has 40GB backup data with about
10 million chunks will need less than 85MB of memory to complete a VM-specific mark-sweep process.
\end{itemize}

### 2013-09-24 snapshot_system/fast14_evaluation.tex
revisions

Figure~\ref{fig:exbin-efficiency-graph2} shows the deduplication efficiency for SRB and VC.
% <<<<<<< HEAD
% Our deduplication efficiency is defined as the percent of duplicate chunks
% which are detected and deduplicated, so only global perfect deduplication can have 100\% efficiency.
% The figure also compares several PDS sizes chosen for VC. ``$\sigma=2\%$'' 
% is defined in Table~\ref{tab:symbol}. 
% =======
Deduplication efficiency is defined as the percent of duplicate chunks
which are detected and deduplicated, so only global perfect deduplication can have 100\% efficiency.
 We also compare several PDS sizes chosen for VC.
%(see Table~\ref{tab:symbol} for definition of ``$\sigma=2\%$'')
%we allocate the space of distributed shared memory  to accommodate 2\%
%of data chunks shared among VMs, as defined in 
With $\sigma=2\%$, memory usage for PDS index lookup per machine is about 100MB per machine
and  the deduplication efficiency can reach over 96.33\%.
When $\sigma=4\%$, the deduplication efficiency can reach 96.9\% while space consumption increases to 200MB per machine. 
%and  the deduplication efficiency can reach 96.489\%.
%When $\sigma=4\%$, the deduplication efficiency can reach 97.027\% while space consumption increases to 200MB per machine. 
The loss of efficiency in VC is caused by the restriction of the physical memory available
in the cluster to support fast in-memory PDS index lookup. 
%In the extreme case if we put all the fingerprint into PDS 
%index ($\sigma=100\%$) then VC will have 100\% efficiency.
SRB can deliver up to 97.79\% deduplication efficiency which is slightly better than VC.
Thus this represents a tradeoff that VC provides better fault tolerance and fast approximate deletion
with a competitive deduplication efficiency.

% SRB can deliver up to 97.935\% deduplication efficiency which is slightly better than VC.
% Thus this represents a tradeoff: VC provides better fault tolerance and fast
% approximate deletion with a competitive (but slightly lower) deduplication
% efficiency.
% >>>>>>> 4dcd68612b052c0f890ed30361a9ed03771b5df1
%its loss of efficiency is caused by the routing of data chunks which restricts the search scope
%of duplicate detection.
%SRB provides slightly better deduplication efficiency than our VC approach, which we accept
%because we are rewarded by architecture-wise benefits such as fault-tolerance and fast 
%Though not our primary goal, this result does show that VC can remove a competitive amount of duplicates.

Figure~\ref{fig:exbin-efficiency-graph2} also shows the effectiveness of the local similarity-based search in VC.
Without this (i.e. when we only check the parent segment at the same offset), when an entire
segment is moved to another location on disk, for example when a file is rewritten
rather than modified in place, the deduplication opportunity is missed. We have found that in
approximately 1/3 of the VMs in our dataset this movement happens frequently,
so the similarity search becomes important for those VMs.
In general, our experiments show that
dirty-bit detection at the segment level can reduce the data size to about 24.14\% of original data, 
which leads about 75.86\% reduction.
Similarity-guided local search can further reduce the data size
to about 12.05\% of original size, namely it delivers additional 12.09\% reduction.
The PDS-based cross-VM detection with $\sigma=2\% $
can reduce the size further to 8.6\% of original size, namely it 
delivers additional 3.45\% reduction.

@@

{\bf Memory and disk bandwidth usage with multi-VM processing}. 
We have further studied the memory and disk bandwidth usage 
when running concurrent VM snapshot backup on each machine with $\delta=2\%$. 
Table ~\ref{tab:resource_usage} gives the resource usage  when
running 1 or multiple  VM backup tasks at the same time on each physical machine. 
``CPU'' column is the percentage of a core usage. 
``Mem'' column includes 100MB memory usage for PDS index and other space cost for executing deduplication tasks such as 
receipt metadata and cache. 
``Read'' column is controlled as 50MB/s bandwidth usage so that other cloud services are not impacted a lot.
``Write'' column is the I/O write usage of QFS and notice that each QFS write leads writes in multiple machines due to
data replication.     
% [ NOT REASONABLE: Notice that memory and CPU used by PDS index and QFS are not included since they belong to 
% the cloud infrastructure services and are shared among many applications.]
%We see our hybrid deduplication scheme only occupies a small amount of system resources. 
%The  local deduplication only needs to keep the parent snapshot recipe and a few similar segment recipes in 
%memory during duplicate detection.

@@

{\bf Processing Time breakdown}.
Figure~\ref{fig:vc_srb_combined} shows
the  average  processing  time of  a VM segment under VC and SRB. 
VC uses $\sigma=2\%$ and $4\%$.
It has a breakdown of time for reading data, updating the metadata, network latency to visit
a remote machine, and index access for fingerprint comparison.
The change of $\sigma$ does not significantly affect the overall backup speed as
PDS lookup takes only a small amount of time.
SRB has a higher index access and fingerprint comparison because once a chunk is routed to a machine,
it relies on this local machine to access its index (often on the disk, supported by a Bloom filter) 
and perform comparison.
VC is faster for index access because it conducts in-memory local search first and then
accesses  the PDS on distributed shared memory.  
SRB spends  slighter more time in  reading data and updates because it also updates the on-disk
local meta data index.
Overall,  VC is faster than SRB, though data reading dominates the processing time for both algorithms.
%Figure~\ref{fig:vc_srb_combined} also reports the average backup time for a VM in VC when
%varying the PDS size.  It lists the time distribution for data reading,
%similarity-guided local search, cross-VM PDS lookup, and non-duplicate data writing. 
%While data reading dominates the backup process, PDS lookup spends about a similar amount
%of time as local search.

>>>>

Figure~\ref{fig:exbin-efficiency-graph2} shows the deduplication efficiency for SRB and VC.
% <<<<<<< HEAD
% Our deduplication efficiency is defined as the percent of duplicate chunks
% which are detected and deduplicated, so only global perfect deduplication can have 100\% efficiency.
% The figure also compares several PDS sizes chosen for VC. ``$\sigma=2\%$'' 
% is defined in Table~\ref{tab:symbol}. 
% =======
Deduplication efficiency is defined as the percent of duplicate chunks
which are detected and deduplicated, so only global perfect deduplication can have 100\% efficiency.
 We also compare several PDS sizes chosen for VC.
%(see Table~\ref{tab:symbol} for definition of ``$\sigma=2\%$'')
%we allocate the space of distributed shared memory  to accommodate 2\%
%of data chunks shared among VMs, as defined in 
With $\sigma=2\%$, memory usage for PDS index lookup per machine is about 100MB per machine
and  the deduplication efficiency can reach 96.489\%.
When $\sigma=4\%$, the deduplication efficiency can reach 97.027\% while space consumption increases to 200MB per machine. 
The loss of efficiency in VC is caused by the restriction of the physical memory available
in the cluster to support fast in-memory PDS index lookup. 
%In the extreme case if we put all the fingerprint into PDS 
%index ($\sigma=100\%$) then VC will have 100\% efficiency.
SRB can deliver up to 97.935\% deduplication efficiency which is slightly better than VC.
Thus this represents a tradeoff: VC provides better fault tolerance and fast
approximate deletion with a competitive (but slightly lower) deduplication
efficiency.
%its loss of efficiency is caused by the routing of data chunks which restricts the search scope
%of duplicate detection.
%SRB provides slightly better deduplication efficiency than our VC approach, which we accept
%because we are rewarded by architecture-wise benefits such as fault-tolerance and fast 
%Though not our primary goal, this result does show that VC can remove a competitive amount of duplicates.

Figure~\ref{fig:exbin-efficiency-graph2} also shows the effectiveness of the local similarity-based search in VC.
Without this (i.e. when we only check the parent segment at the same offset), when an entire
segment is moved to another location on disk, for example when a file is rewritten
rather than modified in place, the deduplication opportunity is missed. We have found that in
approximately 1/3 of the VMs in our dataset this movement happens frequently,
so the similarity search becomes important for those VMs.
In general, our experiments show that
dirty-bit detection at the segment level can reduce the data size to about 24\% of original data, 
which leads  about 76\% reduction.
Similarity-guided local search can further reduce the data size
to about 12\% of original size, namely it delivers additional 12\% reduction.
The PDS-based cross-VM detection with $\sigma=2\% $
can reduce the  size further to 5.6\% of original size, namely it 
delivers additional 6.4\% reduction.

@@

{\bf Memory and disk bandwidth usage with multi-VM processing}. 
We have further studied the memory and disk bandwidth usage 
when running concurrent VM snapshot backup on each machine with $\delta=2\%$. 
Table ~\ref{tab:resource_usage} gives the resource usage  when
running 1 or multiple  VM backup tasks at the same time on each physical machine. 
``CPU'' column is the percentage of a single core used. 
``Mem'' column includes 100MB memory usage for PDS index and other space cost for executing deduplication tasks such as 
receipt metadata and cache. 
``Read'' column is controlled as 50MB/s bandwidth usage so that other cloud services are not impacted a lot.
``Write'' column is the I/O write usage of QFS and notice that each QFS write leads writes in multiple machines due to
data replication.     
% [ NOT REASONABLE: Notice that memory and CPU used by PDS index and QFS are not included since they belong to 
% the cloud infrastructure services and are shared among many applications.]
%We see our hybrid deduplication scheme only occupies a small amount of system resources. 
%The  local deduplication only needs to keep the parent snapshot recipe and a few similar segment recipes in 
%memory during duplicate detection.

@@

{\bf Processing Time breakdown}.
Figure~\ref{fig:vc_srb_combined} shows
the  average  processing  time of  a VM segment under VC and SRB. 
VC uses $\sigma=2\%$ and $4\%$.
It has a breakdown of time for reading data, updating the metadata, network latency to visit
a remote machine, and index access for fingerprint comparison.
The change of $\sigma$ does not significantly affect the overall backup speed as
PDS lookup takes only a small amount of time.
SRB has higher index access and fingerprint comparison costs because all chunk signatures for whom the current
machine does not control the index must be sent over the network, and then rely
on the index holding machine to access its index (often on the disk, supported by a Bloom filter) 
and perform comparison.
VC is faster for index access because it conducts in-memory local search first and then
accesses  the PDS on distributed shared memory, so most chunks can be dealt with locally.  
SRB spends  slighter more time in  reading data and updates because it also updates the on-disk
local meta data index in addition to partition indices.
Overall,  VC is faster than SRB, though data reading dominates the processing time for both algorithms.
%Figure~\ref{fig:vc_srb_combined} also reports the average backup time for a VM in VC when
%varying the PDS size.  It lists the time distribution for data reading,
%similarity-guided local search, cross-VM PDS lookup, and non-duplicate data writing. 
%While data reading dominates the backup process, PDS lookup spends about a similar amount
%of time as local search.
